The promise of community-driven preprints in ecology and evolution

### 2024-10-13 ms/ms.qmd
C3, but also moving around blocks of text
Adjusting C3 and adding more clarity to 3rd paragraph

*EcoEvoRxiv* is one of the few community-driven preprint servers that has paved the way for new initiatives, by accepting multilingual preprints, registered reports, and non-traditional research reports. Such initiatives are distinct from other preprint servers, such as bioRxiv, which only accept empirical research in English. *EcoEvoRxiv* also promotes peer review and community discussion in the hopes of improving the quality of preprints and speeding up their peer-reviewed publication (e.g., Peer Community In -- PCI [@guillemaud2019peer]). For example, we encourage authors to use services such as PCI, which allow for fast, constructive peer review around a preprint with peer reviews being transparent and published online [@guillemaud2019peer].

@@

*EcoEvoRxiv* is run by the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE)[@o2021towards]. Originally launched in 2018 on the Center for Open Science preprint platform, *EcoEvoRxiv* has become a popular preprint server for ecologists and evolutionary biologists. The server has since been adopted by the California Digital Library (CDL). Editors are ecologists and evolutionary biologists from across the globe who volunteer their time to screen papers and push new initiatives in the preprint space. *EcoEvoRxiv* allows authors to submit both preprints and postprints (also known as author-accepted manuscripts). While preprints are versions of manuscripts posted by authors before peer-review, postprints are versions of peer-reviewed and accepted articles but without typesetting and formatting by a journal. The main reason for publishing postprints on a preprint server is to ensure published articles are openly accessible to everyone without a paywall (i.e., green open access). Postprints can be published anytime, provided that journals allow it (which most do; see https://www.sherpa.ac.uk/romeo/).

@@

To better understand preprint (and postprint) use on *EcoEvoRxiv*, we downloaded metadata on the accepted articles available on *EcoEvoRxiv* as of  `r gsub("UTC", "", dates[2])` (see Supplement for more details on methods). We consider both preprints and postprints as ‘articles’. After removing five duplicate titles – suggesting that a few authors uploaded the article as separate submissions rather than updating the existing article – we had data for a total of `r pr` articles with ~55--60 posted approximately monthly in the last two years ([@fig-summary]A).  *EcoEvoRxiv* hosts articles from authors based in `r length(unique(countries$country))` countries, with 90% coming from just `r nrow(most_use)-1` countries. North America, Australia, and European countries upload the most preprints, with many fewer coming from countries in Africa, Central America, and parts of Asia ([@fig-summary]B).  Articles covered all major taxonomic groups, with the most common groups being vertebrates (`r sum_taxa$perc[6]`%), plants (`r sum_taxa$perc[1]`%), and invertebrates (`r sum_taxa$perc[5]`%) ([@fig-summary]C).

>>>>

*EcoEvoRxiv* is one of the few community-driven preprint servers that has paved the way for new initiatives, by accepting multilingual preprints, registered reports, and non-traditional research reports. Such initiatives are distinct from other preprint servers, such as bioRxiv, which only accept empirical research in English. In addition, community-driven servers like EcoEvoRxiv aggregate papers presenting research on similar topics, improving discoverability and opportunities for within-community debate compared with broader preprint servers.  *EcoEvoRxiv* also promotes peer review and community discussion in the hopes of improving the quality of preprints and speeding up their peer-reviewed publication (e.g., Peer Community In -- PCI [@guillemaud2019peer]). For example, we encourage authors to use services such as PCI, which allow for fast, constructive peer review around a preprint with peer reviews being transparent and published online [@guillemaud2019peer]. *EcoEvoRxiv* allows authors to submit both preprints and postprints (also known as author-accepted manuscripts). While preprints are versions of manuscripts posted by authors before peer-review, postprints are versions of peer-reviewed and accepted articles but without typesetting and formatting by a journal. The main reason for publishing postprints on a preprint server is to ensure published articles are openly accessible to everyone without a paywall (i.e., green open access). Postprints can be published anytime, if journals allow it (which most do; see https://www.sherpa.ac.uk/romeo/).

@@

*EcoEvoRxiv* is run by the Society for Open, Reliable, and Transparent Ecology and Evolutionary Biology (SORTEE)[@o2021towards]. Originally launched in 2018 on the Center for Open Science preprint platform, *EcoEvoRxiv* has become a popular preprint server for ecologists and evolutionary biologists. The server has since been adopted by the California Digital Library (CDL). Editors are ecologists and evolutionary biologists from across the globe who volunteer their time to screen papers and push new initiatives in the preprint space. To better understand preprint (and postprint) use on *EcoEvoRxiv*, we downloaded metadata on the accepted articles available on *EcoEvoRxiv* as of  `r gsub("UTC", "", dates[2])` (see Supplement for more details on methods). We consider both preprints and postprints as ‘articles’. After removing five duplicate titles – suggesting that a few authors uploaded the article as separate submissions rather than updating the existing article – we had data for a total of `r pr` articles with ~55--60 posted approximately monthly in the last two years ([@fig-summary]A).

@@

*EcoEvoRxiv* hosts articles from authors based in `r length(unique(countries$country))` countries, with 90% coming from just `r nrow(most_use)-1` countries. North America, Australia, and European countries upload the most preprints, with many fewer coming from countries in Africa, Central America, and parts of Asia ([@fig-summary]B).  Articles covered all major taxonomic groups, with the most common groups being vertebrates (`r sum_taxa$perc[6]`%), plants (`r sum_taxa$perc[1]`%), and invertebrates (`r sum_taxa$perc[5]`%) ([@fig-summary]C).
